Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-25 paper/paper.tex
minor changes

\begin{figure*}[ht]
  \centering
  \includegraphics[trim=0 395 50 0,clip,width=0.65\textwidth]{figs/nids-workflow}
  \caption{\tname\ workflow.}
  \label{fig:overview}
  \vspace{-2ex}
\end{figure*}

@@

The first step of \tname\ is Reverse Engineering. The goal of this
step is to extract options from the malicious traffic. A rule that
contains these options, by construction, captures the malicious
traffic. \tname\ uses two strategies to extract options: 1)~field and
2)~payload. We elaborate these strategies in the following.

For the field strategy, \tname{} parses the input messages looking for
options associated with the data in the fields of a network
message. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from
\suri. Section~\ref{sec:active-recon} shows a concrete example where
the values for options are directly extracted from packet fields. We
constructed the mapping of network fields to options manually, from
the \suri\ documentation.

Unfortunately, the relevant data for the option may be at the message
payload not the fields. Figure~\ref{fig:rule-jboss} shows an example
of a rule that captures traffic containing the string
\CodeIn{"/HtmlAdaptor"} in the payload.  The content strategy focuses
on extracting (content) options from the payload. It is very important
to note that content options are prevalent. For example,
\percRulesWithContent{} of the \numrulessuri{} rules from the public
data set~\cite{emerging-threats-open} we analyzed contain at least one
option of this kind. Figure~\ref{fig:distribution-contents} shows the
histogram of the number of these options per rule for the data set
mentioned above. We limited the number of content options per rule to
10 for space. Rules without \CodeIn{content} options are an exception
and half of the rules contain at least two \CodeIn{content} options.
To find the values associated with these options, \tname{} splits the
payload of the malicious message in tokens using natural language
delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
spaces. We observed, by inspecting existing rules, that
\CodeIn{content} options typically refer to sequence of characters in
the payload of messages separated by these delimiters. This
tokenization step can generate dozens of options.  For example,
considering the example from Section~\ref{sec:content-example},
\tname{} produces a total of \Fix{XXX} content options.

\Mar{Nao esta muito claro como isto funciona. Rodamos sempre as duas
  estrategias (para obter field options e content options)?
  (Precisamos dizer se eh possivel ter regras com os dois tipos de
  opcoes). Como isto funciona 1) quando uma ataque tem varias
  mensagens (e.g., ping scan) e 2) quando temos varios ataques?}

\textbf{Implementation.}\Fix{elaborate...}
, characterized with pcap files~\cite{pcap}, a popular format
to describe network packets.

>>>>

\begin{figure*}[ht]
  \centering
  \includegraphics[trim=0 395 50 0,clip,width=0.65\textwidth]{figs/nids-workflow}
  \caption{The \tname\ workflow.}
  \label{fig:overview}
  \vspace{-2ex}
\end{figure*}

@@

The first step in the \tname's workflow is Reverse Engineering. The
goal of this step is to extract options from the malicious traffic. A
rule that contains these options, by construction, captures the
malicious traffic. \tname\ uses two strategies to extract options:
1)~field (for any protocol) and 2)~payload (for http only). We
elaborate these strategies in the following.

For the field strategy, \tname{} parses the input messages looking for
options associated with the data in the fields of a network
message. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from
\suri. Section~\ref{sec:active-recon} shows a concrete example where
the values for options are directly extracted from packet fields. We
obtained the mapping of network fields to options manually, from the
\suri\ documentation.

Unfortunately, the relevant data for the option may be at the message
payload not the fields. We found empirically that this occurs
frequently in http attacks. Figure~\ref{fig:rule-jboss} shows an
example of a rule that captures traffic containing the string
\CodeIn{"/HtmlAdaptor"} in the payload.  The content strategy focuses
on extracting such kinds of options from the payload. It is very
important to note that content options are prevalent in \suri\ rules
and the majority of these rule prescribe defenses against http
attacks. For example, \percRulesWithContent{} of the \numrulessuri{}
rules from the public data set~\cite{emerging-threats-open} we
analyzed contain at least one content
option. Figure~\ref{fig:distribution-contents} shows the histogram of
the number of these options per rule for the data set mentioned
above. We limited the number of content options per rule to 10 for
space. Furthermore, \Fix{xxx\%} of the rules are for
http.\MyComment{Rules without \CodeIn{content} options are an
  exception and half of the rules contain at least two
  \CodeIn{content} options.}  Consequently, hangling content options
for http is very important.

We observed, by inspecting existing rules, that \CodeIn{content}
options typically refer to sequence of characters in the payload. To
find the values associated with these options, \tname{} splits the
payload of the malicious message in tokens using natural language
delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
spaces. Considering the example from
Section~\ref{sec:content-example}, \tname{} produced a total of 23
content options.

It is worth noting that we found empirically that deciding which
reverse engineering strategy to use based on the protocol was
effective. We used the payload strategy for http and the field
strategy for the other protocols. At the end of this step,
\tname\ produces a seed rule that captures the malicious traffic by
definition.

%% \Mar{Nao esta muito claro como isto funciona. Rodamos sempre as duas
%%   estrategias (para obter field options e content options)?
%%   (Precisamos dizer se eh possivel ter regras com os dois tipos de
%%   opcoes). Como isto funciona 1) quando uma ataque tem varias
%%   mensagens (e.g., ping scan) e 2) quando temos varios ataques?}

%% \textbf{Implementation.}~
%% , characterized with pcap files~\cite{pcap}, a popular format
%% to describe network packets.
